---
title: "Assignment2_1"
author: "wang"
date: "2023-02-11"
output: html_document
---

### Assignment2
### Low Birth Weight
### Analysis1-1: Conduct some basic summary statistics to explore the data, such as the proportion of low birth
weights, proportion of smokers, mean birth weight, etc.


```{r echo=FALSE}
#install.packages("MASS")
library(MASS)
library(ggplot2)
library(ggforce)
library(rstatix)
library(ggpubr)
```

### Compute the proportion of low birth weight baby in this dataset
### Draw the pie plot to visualize the proportion
```{r low weight baby echo=FALSE}
n<-dim(birthwt)[1]
low<-as.data.frame(table(birthwt$low))
levels(low$Var1)<-c("health(>2.5kg)","low weight(<2.5kg)")
pro1<-low[1,2]/n
pro2<-low[2,2]/n
pros1<-sprintf("%1.2f%%",pro1*100)
pros2<-sprintf("%1.2f%%",pro2*100)

low$proportion<-c(pros1,pros2)
low

p1<-ggplot()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(), 
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        legend.title=element_blank(), 
        panel.border = element_blank(),
        panel.background = element_blank())+#去除没用的ggplot背景，坐标轴
  labs(title="Low weight baby proportion")+
  theme(plot.title = element_text(hjust = 0.5))+
  xlab("")+ylab('')+#添加颜色
  scale_fill_manual(values = c('#E5D2DD', '#53A85F'))+
  geom_arc_bar(data=low,
               stat = "pie",
               aes(x0=0,y0=0,r0=0,r=2,
                   amount=Freq,fill=Var1)
  )+#饼图
  annotate("text",x=2.2,y=0,label=pros1,angle=-90)+
  annotate("text",x=-1.6,y=1.5,label=pros2,angle=45)

p1
#fig<-"low"
#name<-paste(paste("./",fig,sep = ""),"pdf",sep = ".")
#ggsave(p1, file=name, width=5.5,height=4)
```


### Compute the proportion of smokers in this dataset
### Draw the pie plot to visualize the proportion
```{r smoke echo=FALSE}
n<-dim(birthwt)[1]
#head(birthwt)
smoker<-as.data.frame(table(birthwt$smoke))
levels(smoker$Var1)<-c("Non-smoker","Smoker")
pro3<-smoker[1,2]/n
pro4<-smoker[2,2]/n
pros3<-sprintf("%1.2f%%",pro3*100)
pros4<-sprintf("%1.2f%%",pro4*100)

smoker$proportion<-c(pros3,pros4)
smoker

p2<-ggplot()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        legend.title=element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank())+#去除没用的ggplot背景，坐标轴
  labs(title="Maternal smoking proportion")+
  theme(plot.title = element_text(hjust = 0.5))+
  xlab("")+ylab('')+#添加颜色
  scale_fill_manual(values = c("#F1BB72","#D6E7A3"))+
  geom_arc_bar(data=smoker,
               stat = "pie",
               aes(x0=0,y0=0,r0=0,r=2,
                   amount=Freq,fill=Var1)
  )+#饼图
  annotate("text",x=2.2,y=0,label=pros3,angle=-90)+
  annotate("text",x=-1.6,y=1.5,label=pros4,angle=45)

p2
#fig<-"smoker"
#name<-paste(paste("./",fig,sep = ""),"pdf",sep = ".")
#ggsave(p2, file=name, width=5.5,height=4)

```

### Compute the proportion of hypertension patients in this dataset
### Draw the pie plot to visualize the proportion
```{r ht}
n<-dim(birthwt)[1]
#head(birthwt)
ht<-as.data.frame(table(birthwt$ht))
levels(ht$Var1)<-c("Health","Hypertension")
pro5<-ht[1,2]/n
pro6<-ht[2,2]/n
pros5<-sprintf("%1.2f%%",pro5*100)
pros6<-sprintf("%1.2f%%",pro6*100)

ht$proportion<-c(pros5,pros6)
ht


p3<-ggplot()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        legend.title=element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank())+#去除没用的ggplot背景，坐标轴
  labs(title="Maternal hypertension proportion")+
  theme(plot.title = element_text(hjust = 0.5))+
  xlab("")+ylab("")+#添加颜色
  scale_fill_manual(values = c("#57C3F3","#E59CC4"))+
  geom_arc_bar(data=ht,
               stat = "pie",
               aes(x0=0,y0=0,r0=0,r=2,
                   amount=Freq,fill=Var1)
  )+#饼图
  annotate("text",x=2.2,y=0,label=pros5,angle=-90)+
  annotate("text",x=-0.6,y=2.1,label=pros6,angle=15)

p3
#fig<-"ht"
#name<-paste(paste("./",fig,sep = ""),"pdf",sep = ".")
#ggsave(p3, file=name, width=5.5,height=4)

```

### Compute the distribution of mothers' age in this dataset
### Draw the density plot to visualize the distribution
```{r}
p4<-ggplot(birthwt, aes(x=age)) + 
     geom_histogram(aes(y=..density..),
                    binwidth=.5,
                    colour="black", fill="white") +
     geom_density(alpha=.2, fill="#FF6666")+
     labs(title="Density distribution of mother age")+
     xlab("age")+ylab("density")+
     theme_classic()+
     theme(plot.title = element_text(hjust = 0.5))
    
p4
```


### Compute the distribution of mothers' weight in this dataset
### Draw the density plot to visualize the distribution
```{r}
p5<-ggplot(birthwt, aes(x=lwt)) +
     geom_histogram(aes(y=..density..),
                    binwidth=.5,
                    colour="black", fill="white") +
     geom_density(alpha=.2, fill="#FF6666")+
     labs(title="Density distribution of mother's weight(pounds)")+
     xlab("weight")+ylab("density")+
     theme_classic()+
       theme(plot.title = element_text(hjust = 0.5))
p5
```
### Compute the some statisic parameters birth weight in this dataset
```{r}
bwtMean<-mean(birthwt$bwt)
bwtMean
max(birthwt$bwt)
min(birthwt$bwt)
median(birthwt$bwt)
```
```{r}
summary<-rbind(low,smoker,ht)
summary
```

### SUMMAY
From the above exploratory dataset,we may conclude these points.
(1)For the baby weight,68.78% baby's weight is considered to be healthy(>2.5kg),while 31.22% babys may suffer from low weight.The average weight of the babys is 2944.567g(>2.5kg);the median weight of the babys is 2977g;the maxinum weight of the babys is 4990g;the minimum weight of the babys is 709g.
(2)For the lifestyle and disease history of the mother,there is 39.15% mothers smokers and 6.35% mothers haing the hypertension.
(3)For the age and weight ,the age of mathers is between 18-25;and the weight of mothers is between 100-130;

### Analysis1-2. Calculate the relative risk of having a low-birth weight infant given the following
```{r}
birthwt[birthwt$age>=20,"ageStatus"]<-"Old"
birthwt[birthwt$age<20,"ageStatus"]<-"Young"
birthwt$ftv<-factor(birthwt$ftv)
levels(birthwt$ftv)<-c("0","1","2","3","4","6")
model <- lm(bwt ~ ., data = birthwt)
summary(model)
```
From the analysis result,we may conclude:
(1) ui is highly related to the baby birth weight;and then is the pregnant race and smoking lifestyle. 


2a. Smoking status
### Use the wilcox.test to compare whether the smoke lifestyles sighificantly affect the baby birth weight.
### The P value is 0.00677,it means if the mother smoke,her baby's weight have the high risk to be lower.
```{r}
birthwt$smoke<-factor(birthwt$smoke)
levels(birthwt$smoke)<-c("Non-smoker","Smoker")
stat.test <- birthwt %>%
  wilcox_test(bwt~smoke)%>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj")%>%
  add_xy_position(x = "smoke",
                  fun="max",
                  step.increase=0.2)
stat.test
p6<-ggplot(birthwt,aes(x=smoke,y=bwt))+
    geom_violin()+
    geom_boxplot(width=0.1)+
    stat_pvalue_manual(stat.test,label = "p.adj.signif")+
    xlab("smoke or not")+ylab("brith weight(g)")+
    theme_classic()
    
p6
```

2b. Hypertension status
### Use the wilcox.test to compare whether the diease history will sighificantly affect the baby birth weight.
### The P value is 0.118,it means the hypertension have no influence to the baby's weight.

```{r}
birthwt$ht<-factor(birthwt$ht)
levels(birthwt$ht)<-c("Health","Hypertension")
stat.test2 <- birthwt %>%
  wilcox_test(bwt~ht)%>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj")%>%
  add_xy_position(x = "ht",
                  fun="max",
                  step.increase=0.2)
stat.test2
p5<-ggplot(birthwt,aes(x=ht,y=bwt))+
    geom_violin()+
    geom_boxplot(width=0.1)+
    stat_pvalue_manual(stat.test2,label = "p.adj.signif")+
    #scale_fill_brewer(palette="Dark2")+
    xlab("hypertension or not")+ylab("brith weight(g)")+
    theme_classic()
    
p5
```


2c. Attending 0 prenatal care visits
### Use the Kruskal-Wallis test to compare whether the fv will sighificantly affect the baby birth weight.
### The P value is 0.323,it means the fv(0,1,2,3,4,6) have no influence to the baby's weight.

```{r}

res.kruskal <- birthwt %>% kruskal_test(ftv ~ bwt)
res.kruskal
```

2d. Giving birth before age 20
### Use the Kruskal-Wallis test to compare whether the mother age will sighificantly affect the baby birth weight.
### The P value is 0.479,it means the mother age have no influence to the baby's weight.

```{r}
res.kruskal2 <- birthwt %>% kruskal_test(ageStatus ~ bwt)
res.kruskal2
```

### Supplement for curiosity:white mother may have the low risk to bore the low weight baby compared to other race? 
```{r}
birthwt$race<-factor(birthwt$race)
levels(birthwt$race)<-c("white","black","other")
res.kruskal0<-birthwt %>%kruskal_test(bwt~race)
res.kruskal0
stat.test0 <- birthwt %>%
   dunn_test(bwt~race,p.adjust.method = "fdr")%>%
   add_significance("p.adj")%>%
   add_xy_position(x = "race",
                   fun="max",
                   step.increase=0.2)
stat.test0
p<-ggplot(birthwt,aes(x=race,y=bwt))+
     geom_violin()+
     geom_boxplot(width=0.1)+
     stat_pvalue_manual(stat.test0,label = "p.adj.signif")+
     #scale_fill_brewer(palette="Dark2")+
     xlab("race")+ylab("brith weight(g)")+
     theme_classic()

p
```

###Analysis1-3. What conclusions can you draw from your analysis?

Conclusion:Apart from some condition that we can not control,such as uterine irritation and race,the lifestyle of the pregnant may be the contributor of the baby's low weight. 



